---
title: "R Notebook"
output: html_notebook
---
```{r}

library(dplyr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(PNWColors)
library(tidyverse)
library(ggtext)
library(here)

here::here()

fig_dir <- here("figs") 
supp_dir <- here("figs", "supp") 

raw_dir <- here("data", "07_indivFunc", "07_raw")  ## raw output from programs like vs2, vibrant, checkv
out_dir <- here("data", "07_indivFunc", "07_output") ## output from R
##default phage hmmm (all (cryptic + intact)
hmm <- read.csv(here("data","03_funcCleaning","03_out", "rd3_hmm_FINAL_with_nophage.csv"), row.names = 1)

hmm <- subset(hmm, hmm$Genus=="Janthinobacterium")

##subset list of all isolates which had NO phage (at all)
nophage <- subset(hmm,hmm$phage_seq=="no_phage")
nophage <- select(nophage, Sequencing.ID, phage_seq, Isolation)
nophage$ignore <- "ignoreme!"
colnames(nophage) <- c("Sequencing.ID2", "phage_seq2", "Isolation2", "ignore")

## remove no_phage hits from hmm 
hmm <- subset(hmm,hmm$phage_seq!="no_phage")

##load bacterial isolate metadata
isos <- read.csv(here("data", "00_metadata", "Isolates.csv"))

#### classify R COG as Unknown
hmm$aux <- ifelse(hmm$deets=="General function (R)", "Unknown (S, No Hits)", hmm$aux)

### set color scheme
auxcolors2 <- c("Unknown (S, No Hits)" ="#b4b4b4", "Phage-associated"="#D4D4D4", "Amino acid (E)" ="#DDFFA0", "Carbohydrate (G)"="#BDEC6F", "Coenzyme (H)"="#97CE2F", "Energy (C)" ="#4E7705", "Information processing (ALBY)"="#A3E4D7",  "Information Processing (ALBY)" = "#A3E4D7", "Inorganic ion (P)"="#43BA8F", "Lipid (I)" ="#148F77", "Membrane (MW)"="#E7F4FF", "Nucleotide (F)"="#BCE1FF", "Phage_moron, auxiliary metabolic gene and host takeover"="#7DCCFF", "Phage takeover or host defense"="#7DCCFF", "Predicted Chitinase" ="#7D3560", "chitinase" ="#7D3560", "Chitinase" ="#7D3560", "Secondary metabolite (Q)"="#098BD9", "Signaling (DTNZU)"="#FCB076", "Transcription (K)"="#F09163", "Translation (JO)"="#C16654", "Weaker predicted chitinase"="#9D654C", "Predicted_AMG"="#FC0FC0", "phosphoadenosine phosphosulfate reductase"="#FC0FC0", "Serine protease"="#FFF200", "Metal protease"="#636363", "ABC Transporters"="#FFFDD0", "No Prophage"="#000000",  "Chitinase, GH18 family" = "#9D654C", "Chitinase, GH19 family"= "#7D3560") 

### set order on legend
desired_order <- c("ABC Transporters", "Metal protease", "Serine protease", "phosphoadenosine phosphosulfate reductase", "Chitinase", "chitinase", "Predicted Chitinase", "Weaker predicted chitinase", "Chitinase, GH18 family", "Chitinase, GH19 family", "Translation (JO)", "Transcription (K)", "Signaling (DTNZU)", "Defense (V)", "Secondary metabolite (Q)", "Nucleotide (F)", "Membrane (MW)", "Lipid (I)", "Inorganic ion (P)", "Information Processing (ALBY)", "Information processing (ALBY)", "Energy (C)", "Coenzyme (H)", "Carbohydrate (G)", "Amino acid (E)", "Predicted_AMG", "Phage takeover or host defense", "Phage-associated", "Unknown (S, No Hits)", "No Prophage")


##set fig theme
theme2 <- theme_light() +
  theme(
      legend.title = element_blank(),
      legend.position="bottom",
      legend.text = element_text(size=15),
      plot.title = element_text(size=15),
      axis.text.x = element_text(size=15),
      axis.text.y = element_text(size=15),
      axis.title.y = element_text(size=15),
      axis.title.x = element_text(size=15),
      axis.ticks.length = unit(.3, "cm"),
      plot.margin=unit(c(1,1,1,1),"cm"),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(color = "black", size = 15),)
```
###all FACET


## INDIV Isos

```{r}

hmm.indiv <- hmm
###create tally
hmm.indiv$tally <- 1


###Calculate functional distribution for INDIVIDUAL phages

###count number of ORFS per phage
prot.phage <- hmm.indiv %>%
  group_by(phage_seq, intact_est2) %>%
  summarise(TotalORF.perPhage = sum(tally))

### count number of ORFs per phage by functional category (+ keep track of Isolation, intact estimates)
indiv.phage <- hmm.indiv %>%
  group_by(aux, aux2, COG.broad, comp, comp2,, deets, xtremedeets, phrog_category, nogname, Nog.abbrev, Isolation, Sequencing.ID, phage_seq, intact_est2) %>%
  summarise(new.freq = sum(tally))

##combine count data from above
indiv.phage.freq <- merge(prot.phage, indiv.phage, by="phage_seq", all=TRUE)

###remove any categories with NA for now 
indiv.phage.freq <- subset(indiv.phage.freq, indiv.phage.freq$phage_seq!="NA")

### calc the percentage a category takes up that portin of the phage genome (e.g. total genome is 20 genomes, 2 of which are carbon metabolism (2/20 = 10% of genome))
indiv.phage.freq$perc.fq <- indiv.phage.freq$new.freq/indiv.phage.freq$TotalORF.perPhage*100

### add bacterial isolates that did not have any phages back into dataset and set all categories to 0 / uses the columns ending in 2 from the no_phage set to repopulate the original columns
indiv.phage.freq <- merge(indiv.phage.freq, nophage, by.x="phage_seq", by.y="phage_seq2", all=TRUE)
indiv.phage.freq$perc.fq[is.na(indiv.phage.freq$perc.fq)] <- 0
indiv.phage.freq$new.freq[is.na(indiv.phage.freq$new.freq)] <- 0
indiv.phage.freq[is.na(indiv.phage.freq)] <- "NA"

indiv.phage.freq$Isolation <- ifelse(indiv.phage.freq$ignore=="ignoreme!", indiv.phage.freq$Isolation2, indiv.phage.freq$Isolation)

phage.func.indiv <- indiv.phage.freq

###Calculate functional distribution for all phages of a Bacterial host

### count number of ORFS per bacterial host
prot.iso <- hmm.indiv %>%
  group_by(Sequencing.ID, intact_est2) %>%
  summarise(TotalORF.perIso = sum(tally))

prot.iso <- unite(prot.iso, iso_phage, Sequencing.ID, intact_est2, sep = "_", remove=TRUE)


### count number of hits to a Category per bacterial host
indiv.iso <- hmm.indiv %>%
  group_by(aux, aux2, COG.broad, comp, comp2, deets, xtremedeets, phrog_category, nogname, Nog.abbrev, Isolation, Sequencing.ID, intact_est2) %>%
  summarise(new.freq = sum(tally))

indiv.iso <- unite(indiv.iso, iso_phage, Sequencing.ID, intact_est2, sep = "_", remove=FALSE)


### merge count data
indiv.iso.freq <- merge(prot.iso, indiv.iso, by="iso_phage", all=TRUE)
indiv.iso.freq <- subset(indiv.iso.freq, indiv.iso.freq$Sequencing.ID!="NA")

### find percentage of ORF Category that contributes to a bacterial host's "phage pangenome"
indiv.iso.freq$perc.fq <- indiv.iso.freq$new.freq/indiv.iso.freq$TotalORF.perIso*100

## add in no isolates again
indiv.iso.freq <- merge(indiv.iso.freq, nophage, by.x="Sequencing.ID", by.y="Sequencing.ID2", all=TRUE)
indiv.iso.freq$perc.fq[is.na(indiv.iso.freq$perc.fq)] <- 0
indiv.iso.freq$new.freq[is.na(indiv.iso.freq$new.freq)] <- 0
indiv.iso.freq[is.na(indiv.iso.freq)] <- "NA"

indiv.iso.freq$Isolation <- ifelse(indiv.iso.freq$ignore=="ignoreme!", indiv.iso.freq$Isolation2, indiv.iso.freq$Isolation)

phage.func.iso <- indiv.iso.freq
```

```{r}
phage.func.iso$facet1 <- "Genes"
### for isos with no phage, set the "intact" status to degraded so they are not lost when graphing
phage.func.iso$intact_est2 <- ifelse(phage.func.iso$intact_est2=="NA", "Degraded", phage.func.iso$intact_est2)


aux.iso.meta <- subset(phage.func.iso, phage.func.iso$aux!="Unknown (S, No Hits)" & phage.func.iso$aux!="Phage-associated")
aux.iso.meta$facet1 <- "Auxillary Genes"

iso.func <- rbind(phage.func.iso, aux.iso.meta)

```


###MEGA GRAPH by Genus host


```{r}
## make labels on graphs nicer
iso.func$Isolation.clean <- ifelse(iso.func$Isolation=="Enviro", "Environment", iso.func$Isolation)
iso.func$Isolation.clean <- ifelse(iso.func$Isolation=="Host_non", "Host (other)", iso.func$Isolation.clean)
iso.func$Isolation.clean <- ifelse(iso.func$Isolation=="Amphibian", "Amphibian Host", iso.func$Isolation.clean)


iso.func <- unite(iso.func, label, Isolation.clean, Sequencing.ID, sep = ", ", remove=FALSE)
iso.func$label_word <-"Prophage"
iso.func <- unite(iso.func, intact_nice, intact_est2, label_word, sep = " ", remove=FALSE)
iso.func <- unite(iso.func, facet_group, intact_nice, facet1,  sep = " ", remove=FALSE)


##creates invisible placeholder in graph for isolates w/ no phage
iso.func$perc.fq <- ifelse(iso.func$aux=="NA", 0.0, iso.func$perc.fq)

## label letters are for composite figure
iso.func$facet_group <- factor(iso.func$facet_group,
  levels = c('Intact Prophage Genes', 'Intact Prophage Auxillary Genes', 'Degraded Prophage Genes', 'Degraded Prophage Auxillary Genes'),
  labels = c( 'C. Intact Genes', 'D. Intact Auxillary Genes', 'E. Degraded Genes', 'F. Degraded Auxillary Genes'))

gg.isofunc.jan <- ggplot(data=iso.func, aes(fill=aux, x = label, y=perc.fq)) + geom_bar(stat="identity")+ coord_flip() +xlab("")+ylab("Percent of Total Prophage Genes")+theme2 + scale_fill_manual(values = auxcolors2) + guides(fill=guide_legend(ncol=6))+ theme(axis.text.y = element_text(size=16)) + facet_wrap(~ facet_group, ncol = 4, scales = "free_x")+ theme(legend.text = element_text(size=12), strip.text = element_text(size = 13))

gg.isofunc.jan




## save gg image to be used in composite figure later (see 09_.rmd)
#saveRDS(gg.isofunc.jan, file="data/07_indivFunc/07_out/gg_isofunc.rds")
## save RDS was not working well with here() package, regular path used instead.        
#ggsave(here(fig_dir, "F3.pdf"), width=15, height=11)

```

```{r}
### to summarize hits
phage.func.iso <- subset(phage.func.iso, phage.func.iso$iso_phage!="NA")

##aux category used to calc descp statistics in paper, as that is what the figure represents

##METABOLISM aux (antifungals in own categories) summary
aux.summ <- phage.func.iso %>%
  group_by(aux, intact_est2) %>%
  summarise(cat.hits.per.genera = sum(new.freq), .groups = "drop") %>%
  pivot_wider(
    names_from = intact_est2,
    values_from = cat.hits.per.genera,
    values_fill = 0
  )
a.intsum <- sum(aux.summ$Intact) #1747
a.degsum <- sum(aux.summ$Degraded) #479
aux.summ$int.perc <- aux.summ$Intact/a.intsum * 100
aux.summ$deg.perc <- aux.summ$Degraded/a.degsum * 100



aux.summ$perc <- aux.summ$cat.hits.per.genera/sum.hits*100

##METABOLISM aux2 (antifungals in original categories) summary
aux2.summ <- phage.func.iso %>%
  group_by( aux2) %>%
  summarise(cat.hits.per.genera = sum(new.freq))
aux2.summ$perc <- aux2.summ$cat.hits.per.genera/sum.hits*100


##METABOLISM summary
metabcomp.summ <- phage.func.iso %>%
  group_by(comp) %>%
  summarise(cat.hits.per.genera = sum(new.freq))
metabcomp.summ$perc <- metabcomp.summ$cat.hits.per.genera/sum.hits*100


##METABOLISM summary
metabcomp2.summ <- phage.func.iso %>%
  group_by(comp2) %>%
  summarise(cat.hits.per.genera = sum(new.freq))
metabcomp2.summ$perc <- metabcomp2.summ$cat.hits.per.genera/sum.hits*100

```
### 2 by 2 of visualization

iso.func$facet_group <- factor(iso.func$facet_group, levels = c('Genes of Degraded Prophages', 'Genes of Intact Prophages','Aux Genes of Degraded Prophages', 'Aux Genes of Intact Prophages'))


#iso.func$aux <- ifelse(iso.func$aux=="NA", "Unknown (S, No Hits)", iso.func$aux)


gg.isofunc2 <- ggplot(data=iso.func, aes(fill=aux, x = label, y=perc.fq)) + geom_bar(stat="identity") + coord_flip() +xlab("")+ylab("Percent of Prophage Genes") +theme2 + scale_fill_manual(values = auxcolors2) + guides(fill=guide_legend(ncol=1))+ theme(axis.text.y = element_markdown()) + facet_wrap(~ facet_group, ncol = 2, scales = "free_x") + theme(legend.position = "right", legend.text = element_text(size=10), axis.text = element_text(size=10))


gg.isofunc2




